|

### 2023-05-16 paper/01-introduction.Rmd
Make some small edits to wording

In recent years, tracking data have replaced traditional box-score statistics and play-by-play data as the state of the art in sports analytics.
Numerous sports are collecting and releasing data on player and ball locations on the playing surface over the course of a game.
This spatio-temporal multi-resolution source of data has provided exceptional opportunities for researchers to perform advanced studies at a more granular level to deepen our understanding of different sports.
For complete surveys on how tracking data have transformed sports analytics, see @Macdonald2020Recreating, @Baumer2023Big, and @Kovalchik2023Player.

@@

Our focus of this manuscript is specifically on measuring the performance of defensive linemen in the NFL. There are two main types of defensive linemen in American football: (1) defensive tackles (DTs) located within the interior of the line and (2) defensive ends (DEs) located along the edges (see Figure \ref{fig:formation} (top) for example formation with DTs and DEs). The primary purpose of both positions is to rush the QB on passing plays, with DEs displaying superiority in observed pass rushing ability [@Eager2018nfl]. Additionally, within DTs there are nose tackles (NTs) who directly line up across from the ball at the line of scrimmage (see Figure \ref{fig:formation} (bottom) for example defensive formation with a NT). Typically, teams will employ either one NT or two DTs on the interior with DEs along either side of the defensive line. In addition to defensive lineman, other positions may attempt to rush the QB on blitzing plays such as outside linebackers (OLBs), interior linebackers (ILBs), and potentially members of the secondary (cornerback, free safety, strong safety) who typically play pass coverage.

@@

The remainder of this manuscript is outlined as follows.
We first describe the player tracking data provided by the Big Data Bowl 2023 in Section \ref{sec:data}.
We then introduce the mathematical motivation and definition of our measure STRAIN and our modeling approach in Section \ref{sec:methods}.
Next, we present applications of STRAIN and study different statistical properties of the metric in Section \ref{sec:results}.
We close with our discussion of future directions related to this work in Section \ref{sec:discussion}.

>>>>

In recent years, tracking data have replaced traditional box-score statistics and play-by-play data as the state of the art in sports analytics.
Numerous sports are collecting and releasing data on player and ball locations on the playing surface over the course of a game.
This multi-resolution, spatio-temporal source of data has provided exceptional opportunities for researchers to perform advanced studies at a more granular level to deepen our understanding of different sports.
For complete surveys on how tracking data have transformed sports analytics, see @Macdonald2020Recreating, @Baumer2023Big, and @Kovalchik2023Player.

@@

Our focus of this manuscript is specifically on measuring the performance of defensive linemen in the NFL. There are two main types of defensive linemen in American football: (1) defensive tackles (DTs) located within the interior of the line and (2) defensive ends (DEs) located along the edges; see Figure \ref{fig:formation} (top) for example formation with DTs and DEs. The primary purpose of both positions is to rush the QB on passing plays, with DEs displaying superiority in observed pass rushing ability [@Eager2018nfl]. Additionally, within DTs there are nose tackles (NTs) who directly line up across from the ball at the line of scrimmage; see Figure \ref{fig:formation} (bottom) for example defensive formation with a NT. Typically, teams will employ either one NT or two DTs on the interior with DEs along either side of the defensive line. In addition to defensive lineman, other positions may attempt to rush the QB on blitzing plays such as outside linebackers (OLBs), interior linebackers (ILBs), and potentially members of the secondary (cornerback, free safety, strong safety) who typically play pass coverage.

@@

The remainder of this manuscript is outlined as follows.
We first describe the player tracking data provided by the Big Data Bowl 2023 in Section \ref{sec:data}.
We then introduce the mathematical motivation and definition of our measure STRAIN, followed by our modeling approach in Section \ref{sec:methods}.
Next, we present applications of STRAIN and study different statistical properties of the metric in Section \ref{sec:results}.
We close with our discussion of future directions related to this work in Section \ref{sec:discussion}.

### 2023-05-16 paper/02-data.Rmd
Make some small edits to wording

In the forthcoming analysis, we rely on the data from the NFL Big Data Bowl 2023 provided by the NFL Next Gen Stats tracking system.
The data corresponds to 8,557 passing plays across 122 games in the first eight weeks of the 2021 NFL regular season. 
For each play, we have information on the on-field location, speed, angle, direction, and orientation of each player on the field and the football at a rate of 10 Hz (i.e. 10 measurements per second), along with event annotations for each frame such as ball snap, pass forward, and quarterback sack, to name a few.

@@

Along with the tracking information, the 2023 BDB includes scouting data provided by Pro Football Focus (PFF). 
This contains manually-collected player-level information, such as the player's role (e.g., whether they are a pass rusher and pass blocker) and credited events (e.g., player is credited with hitting the QB on the play).
In this manuscript, we use the PFF data to identify 36,362 unique pass rush attempts by players designated in the "pass rush" role across all plays.
For context, Figure \ref{fig:pass_rushers_count} displays the distribution of the number of observed pass rushers involved in a play, ranging from 1 to 8 with 4 pass rushers (i.e. 4-man rush formation) as the most common value.
We also leverage the PFF scouting data to count how many hits, hurries, and sacks each pass rusher is credited with across the span of observed data. 
Additionally, we use the PFF player roles to identify the blocking matchup for each pass rusher in order to adjust for opponent strength, as discussed in Section \ref{sec:methods}.

>>>>

In the forthcoming analysis, we rely on the data from the NFL Big Data Bowl 2023 provided by the NFL Next Gen Stats tracking system.
The data corresponds to 8,557 passing plays across 122 games in the first eight weeks of the 2021 NFL regular season. 
For each play, we have information on the on-field location, speed, angle, direction, and orientation of each player on the field and the football at a rate of 10 Hz (i.e., 10 measurements per second), along with event annotations for each frame such as ball snap, pass forward, and quarterback sack, to name a few.

@@

Along with the tracking information, the 2023 BDB includes scouting data provided by Pro Football Focus (PFF). 
This contains manually-collected player-level information, such as the player's role (e.g., whether they are a pass rusher and pass blocker) and credited events (e.g., player is credited with hitting the QB on the play).
In this manuscript, we use the PFF data to identify 36,362 unique pass rush attempts by players designated in the "pass rush" role across all plays.
For context, Figure \ref{fig:pass_rushers_count} displays the distribution of the number of observed pass rushers involved in a play, ranging from 1 to 8 with 4 pass rushers (i.e., 4-man rush formation) as the most common value.
We also leverage the PFF scouting data to count how many hits, hurries, and sacks each pass rusher is credited with across the span of observed data. 
Additionally, we use the PFF player roles to identify the blocking matchup for each pass rusher in order to adjust for opponent strength, as discussed in Section \ref{sec:methods}.

### 2023-05-16 paper/03-methods.Rmd
Make some small edits to wording

To this end, we fit a multilevel model to evaluate pass rushers' impact on the average STRAIN observed in a play, while accounting for their pass blocker, QB, and play-level attributes.
We identify the pass blocker linked with the pass rusher of interest using the data provided by PFF.
Since there can be multiple blockers matching up with a rusher, for simplicity, we consider the nearest blocker positioned to the pass rusher at the start of the play.
We use random intercepts for the three player groups: (1) pass rushers as $R$, (2) nearest pass blockers as $B$, and (3) QBs as $Q$.
The coefficients $\boldsymbol \beta$ for covariate vector $\mathbf{x_{ij}}$ for pass rusher $j$ in play $i$ are estimated as fixed-effects. 
Our model for the average STRAIN by pass rusher $j$ on play $i$ is as follows.
\begin{equation}
\begin{aligned}
\overline{\text{STRAIN}}_{ij} &\sim N(R_{j[i]} + B_{b[ij]} + Q_{q[i]} + \mathbf{x_{ij}} \cdot \boldsymbol{\beta}, \sigma^2), \text{ for } i = 1, \dots, n \text{ plays} \\
R_{j} &\sim N(\mu_R, \sigma^2_R), \text{ for } j = 1, \cdots, \text{ \# of pass rushers}, \\
B_{b} &\sim N(\mu_B, \sigma^2_B), \text{ for } b = 1, \cdots, \text{ \# of pass blockers}, \\
Q_{q} &\sim N(\mu_Q, \sigma^2_Q), \text{ for } q = 1, \cdots, \text{ \# of QBs}.
\end{aligned}
\end{equation}

>>>>

To this end, we fit a multilevel model to evaluate pass rushers' impact on the average STRAIN observed in a play, while accounting for their pass blocker, QB, and play-level attributes.
We identify the pass blocker linked with the pass rusher of interest using the data provided by PFF.
Since there can be multiple blockers matching up with a rusher, for simplicity, we consider the nearest blocker positioned to the pass rusher at the start of the play.
We use random intercepts for the three player groups: (1) pass rushers as $R$, (2) nearest pass blockers as $B$, and (3) QBs as $Q$.
The coefficients $\boldsymbol \beta$ for covariate vector $\mathbf{x_{ij}}$ for pass rusher $j$ in play $i$ are estimated as fixed-effects. 
Our model for the average STRAIN by pass rusher $j$ on play $i$ is as follows.
\begin{equation}
\begin{aligned}
\overline{\text{STRAIN}}_{ij} &\sim N(R_{j[i]} + B_{b[ij]} + Q_{q[i]} + \mathbf{x_{ij}} \boldsymbol{\beta}, \sigma^2), \text{ for } i = 1, \dots, n \text{ plays} \\
R_{j} &\sim N(\mu_R, \sigma^2_R), \text{ for } j = 1, \cdots, \text{ \# of pass rushers}, \\
B_{b} &\sim N(\mu_B, \sigma^2_B), \text{ for } b = 1, \cdots, \text{ \# of pass blockers}, \\
Q_{q} &\sim N(\mu_Q, \sigma^2_Q), \text{ for } q = 1, \cdots, \text{ \# of QBs}.
\end{aligned}
\end{equation}

### 2023-05-16 paper/04-results.Rmd
Make some small edits to wording

In addition to adjusting for opposition, our multilevel model allows us to provide some notion of uncertainty about the pass rusher's effect on the play level STRAIN.
Figure \ref{fig:fig_rankings_boot} displays the mean varying intercepts for the top ten pass rushers for each position along with the corresponding 95% confidence intervals. 
We observe similar rankings based on the random intercepts in comparison to the average STRAIN results in Tables \ref{tab:edge} and \ref{tab:interior}, with Myles Garrett (DE), Aaron Donald (DT), Vita Vea (NT), and T.J. Watt (OLB) as the first-ranked pass rusher for their respective position. 
We observe different levels of variability across these top players, with only a subset of players with intervals strictly above zero. This is not necessarily surprising given the limited sample of data.
Additionally, we are ignoring correlated performance between defensive linemen on the same team.
For instance, the top two defensive ends, Myles Garrett and Takkarist McKinley (see the top panel of Figure \ref{fig:fig_rankings_boot}), are teammates on the Cleveland Browns.
McKinley is a surprising name on our list and we suspect that his high rank is mostly due to playing on the same defensive unit with Garrett, who is well-known as an elite pass rusher.
We leave the exploration of adjusting for teammates for future work.

>>>>

In addition to adjusting for opposition, our multilevel model allows us to provide some notion of uncertainty about the pass rusher's effect on the play level STRAIN.
Figure \ref{fig:fig_rankings_boot} displays the mean varying intercepts for the top ten pass rushers for each position along with the corresponding 95% confidence intervals. 
We observe similar rankings based on the random intercepts in comparison to the average STRAIN results in Tables \ref{tab:edge} and \ref{tab:interior}, with Myles Garrett (DE), Aaron Donald (DT), Vita Vea (NT), and T.J. Watt (OLB) as the first-ranked pass rusher for their respective position. 
We observe different levels of variability across these top players, with only a subset of players with intervals strictly above zero. This is not necessarily surprising given the limited sample of data.
Additionally, we are ignoring correlated performance between defensive linemen on the same team.
For instance, the top two defensive ends, Myles Garrett and Takkarist McKinley (see the top panel of Figure \ref{fig:fig_rankings_boot}), were teammates on the Cleveland Browns during this period of time.
McKinley is a surprising name on our list and we suspect that his high rank is mostly due to playing on the same defensive unit with Garrett, who is well-known as an elite pass rusher.
We leave the exploration of adjusting for teammates for future work.
